---
title: "Olympic History EDA"
author: "By Abiodun Onadeji"
date: "2022-11-10"
output: html_document
---
# **Olympic History Exploratory Data Analysis with R**

First and foremost, let us load our library that we would be working with in this Analysis
```{r}
#Loading Library

library(tidyverse)
library(skimr)
library(janitor)

```
Lets import our CSV Data set. We would name the athlete_event and noc_region CSV file to Event and Region respectively

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#Importing the CSV Files

Event <- read_csv("athlete_events.csv") 

Region <- read_csv("noc_regions.csv")
```

Let us quickly have a sense of our variables and data frame

```{r}
# Variables of Event Data frame

colnames(Event)

```

```{r}
# Variables of Region Data frame

colnames(Region)

```

```{r}
# overview of Event Data frame

glimpse(Event)

```
The Event data have  271,116 observations and 15 variables

```{r}
# overview of Region Data frame

glimpse(Region)

```

The Region data have  230 observations and 3 variables
-------------------------------------------------------------
Before cleaning our data, let us join the Event & Region Data frame Using the "NOC" Variable and storing in the object "Event_Region" as the new Data frame

```{r}
# joining the Event and Region Data frame by "NOC"

Event_Region <- Event %>% 
  inner_join(Region, by = "NOC")

```

# **Data Cleaning**

```{r}
#Variable names

colnames(Event_Region)

```

```{r}
#cleaning the variable names

Event_Region <- clean_names(Event_Region)

```


```{r}
# variable type

sapply(Event_Region, class)

```

```{r}
# Droping the notes Variable

Event_Region$notes <- NULL
names(Event_Region)

```

```{r}
#summary of the data frame

summary(Event_Region)

```

Lets us convert the variables sex, season and medal to factor from character since they are categorical variables

```{r}
#sex character variable to factor variable

Event_Region$sex <- as.factor(Event_Region$sex)

```


```{r}
#season character variable to factor variable

Event_Region$season <- as.factor(Event_Region$season)

```


```{r}
#medal character variable to factor variable

Event_Region$medal <- as.factor(Event_Region$medal)

```

let's verify that our variables have been converted successfully

```{r}
#Variable type

sapply(Event_Region, class)

```

We are going to be answering the following 25 questions in our Exploratory Data Analysis of the Olympic History data set.

# **Questions**

1. How many Olympic games have taken place?
2. List all the Olympic games along with the cities where they were held.
3. List every region that took part in the 1896 Summer Games.
4. Mention the total number of countries that competed in each Olympic game.
5. Which Olympic game has the most and least countries participating?
6. Which nation or country has competed in each and every Olympic Game?
7. Get the top 5 medal-winning nations from the Olympics. Success is measured by the total number of medals earned.
8. How many gold, silver, and bronze medals did each nation win in total? A nation must take home at least one of each medal.
9. List all of Nigeria's gold, silver, and bronze medals from each Olympic game.
10. List the total number of gold, silver, and bronze medals each nation has earned at each Olympic game.
11. Which country has never won a gold medal but has instead received a silver or bronze?
12. Which sport has Nigerian athletes won the most medals?
13. Analyze each Olympic games in which Nigeria won a football medal, noting the total number of medals obtained.
14. Name the sport that was played at every summer Olympic Games.
15. Which Olympic game have only ever been played once?
16. How many different sports are featured in each Olympic game?
17. Which gender took part in the Olympic games the most?
18. What is the Total Number of Participants Over the Year?
19. What was the gender participation breakdown throughout the course of the year?

20. What were the Medals totaling gold, silver, and bronze received by each gender?
21. List the gold medalists that are under 15 years old.
22. Who are the oldest athletes to have won a gold medal?
23. Who are the top 10 athletes who have won the most gold medals?
24. Get the top 10 athletes by number of medals earned (gold, silver, or bronze).
25. What age group participated in the game the most?

**Question 1:**
How many Olympic games have taken place?

```{r}
#Question 1

n_distinct(Event_Region$games)

```

**Question 2:**
List all the Olympic games along with the cities where they were held.

```{r}
#Question 2

Event_Region %>% 
  select(games, city) %>%
  distinct() %>% 
  arrange(games)

```
The first Olympic game ever was in 1896 Summer which was
hosted in Athina city

**Question 3**
 List every region that took part in the 1896 Summer Games.
 
```{r}
#Question 3

Event_Region %>% 
  select(games, region) %>% 
  distinct() %>% 
  filter(games=="1896 Summer")
```
Twelve (12) Region participated in 1896 Summer game among which were
Greece, Uk, Switzerland etc

**Question 4:**
Mention the total number of countries that competed in each Olympic game.
```{r}
#Question 4

Event_Region %>% 
  select(games, total_countries = region) %>% 
  distinct() %>% 
  aggregate(total_countries~games, FUN = length)
```


**Question 5:**
Which Olympic game has the most and least countries participating?
```{r}
#Question 5

Event_Region %>% 
  select(games, region) %>% 
  distinct() %>% 
  aggregate(region ~ games, FUN = length) %>% 
  summarise(lowest_country_partcipation  = c(min(games),min(region)),
            highest_country_partcipation= c(max(games),max(region)))
```


**Question 6:**
Which nation or country has competed in each and every Olympic Game?
```{r}
#Question 6

Event_Region %>% 
  select(region, games) %>%
  distinct() %>% 
  aggregate(games~region, FUN = length) %>% 
  filter(games == 51)
```

**Question 7:**
Get the top 5 medal-winning nations from the Olympics. Success is measured by the total number of medals earned.
```{r}
#Question 7

Event_Region %>% 
  select(country = region, total_medals = medal) %>% 
  aggregate(total_medals ~ country, FUN = length) %>% 
  arrange(desc(total_medals)) %>% 
  slice(1:5)
```


**Question 8:**
How many gold, silver, and bronze medals did each nation win in total? A nation must take home at least one of each medal.
```{r}
#Question 8

Event_Region %>% 
  group_by(Country = region, medal) %>% 
  filter(medal %in% c("Gold", "Silver", "Bronze")) %>% 
  summarise(n=n()) %>% 
  pivot_wider(names_from = "medal", values_from = "n") %>% 
  select(Gold, Silver, Bronze) %>% 
  na.omit() %>% 
  arrange(desc(Gold))
```


**Question 9:**
List all of Nigeria's gold, silver, and bronze medals from each Olympic game.
```{r}
#Question 9

Event_Region %>% 
  group_by(games, Country = region, medal) %>% 
  filter(medal %in% c("Gold", "Silver", "Bronze")) %>% 
  summarise(n=n()) %>% 
  filter(Country == "Nigeria") %>% 
  pivot_wider(names_from = "medal", values_from = "n")%>% 
  select(Gold, Silver, Bronze) %>% 
  mutate(Gold = replace_na(Gold, 0), 
         Silver= replace_na(Silver, 0),
         Bronze = replace_na(Bronze, 0)) %>% 
  arrange(desc(Gold))
```

**Question 10:**
List the total number of gold, silver, and bronze medals each nation has earned at each Olympic game.
```{r}
#Question 10

Event_Region %>%
  group_by(games, region, medal) %>% 
  summarise(n=n()) %>% 
  na.omit() %>% 
  pivot_wider(names_from = "medal", values_from = "n") %>% 
  select(Gold, Silver, Bronze) %>%
  mutate(Gold = replace_na(Gold, 0), 
         Silver= replace_na(Silver, 0),
         Bronze = replace_na(Bronze, 0))
```


**Question 11:**
Which country has never won a gold medal but has instead received a silver or bronze?
```{r}
#Question 11

Event_Region %>% 
  group_by(region, medal) %>% 
  summarise(n=n()) %>% 
  na.omit() %>% 
  pivot_wider(names_from = "medal", values_from = "n")%>%
  select(Gold, Silver, Bronze) %>%
  mutate(Gold = replace_na(Gold, 0), 
         Silver= replace_na(Silver, 0),
         Bronze = replace_na(Bronze, 0)) %>% 
  filter(Gold == 0 & Silver > 0 & Bronze > 0)
```

**Question 12:**
Which sport has Nigerian athletes won the most medals?
```{r}
#Question 12

Event_Region %>% 
  select(region, sport, medal) %>% 
  aggregate(medal ~ region + sport , FUN = length) %>% 
  filter(region == "Nigeria") %>% 
  arrange(desc(medal))
```


**Question 13:**
Analyze each Olympic games in which Nigeria won a football medal, noting the total number of medals obtained.
```{r}
#Question 13

Event_Region %>% 
  select(team, sport, games, medal) %>% 
  aggregate(medal ~ team + sport + games , FUN = length) %>% 
  filter(team == "Nigeria" & sport == "Football") %>% 
  arrange(desc(medal))
```


**Question 14:**
Name the sport that was played at every summer Olympic Games.
```{r}
#Question 14

Event_Region %>% 
  select(sport, season) %>% 
  distinct() %>% 
  filter(season == "Summer")
```

**Question 15:**
Which Olympic game have only ever been played once?
```{r}
#Question 15

Event_Region %>% 
  select(sport, no_of_games = games) %>% 
  distinct() %>% 
  aggregate(no_of_games~sport, FUN = length) %>% 
  filter(no_of_games == 1)
```

**Question 16:**
How many different sports are featured in each Olympic game?
```{r}
#Question 16

Event_Region %>% 
  select(games, sport) %>% 
  distinct() %>% 
  aggregate(sport ~ games, FUN = length) %>% 
  arrange(desc(sport))
```

**Question 17:**
Which gender took part in the Olympic games the most?
```{r}
#Question 17

Event_Region %>% 
  group_by(sex) %>% 
  summarise(total=n())
```
Male are more involved

**Question 18:**
What is the Total Number of Participants Over the Year?
```{r}
#Question 18

Event_Region %>% 
  select(year, season, total_participant = sex) %>% 
  aggregate(total_participant ~ year + season, FUN = length) %>% 
  arrange(year)
```


**Question 19:**
What was the gender participation breakdown throughout the course of the year?
```{r}
#Question 19

Event_Region %>%
  group_by(year, season, sex) %>% 
  summarise(n=n()) %>% 
  pivot_wider(names_from = "sex", values_from = "n")%>%
  select(male = M, female =F) %>% 
  mutate(male = replace_na(male, 0),
         female = replace_na(female,0))
```

**Question 20:**
What were the Medals totaling gold, silver, and bronze received by each gender?
```{r}
#Question 20

Event_Region %>% 
  group_by(sex, medal) %>% 
  summarise(n=n()) %>% 
  drop_na() %>%
  pivot_wider(names_from = "medal", values_from = "n") %>% 
  select(Gold, Silver, Bronze) %>% 
  mutate(total = (Gold + Silver + Bronze))
```

**Question 21:**
List the gold medalists that are under 15 years old.
```{r}
#Question 21

Event_Region %>% 
  select(name, age, region, medal) %>% 
  filter(age < 15 & medal == "Gold") %>%
  arrange(region, age) %>% 
  distinct()
```

**Question 22:**
Who are the oldest athletes to have won a gold medal?
```{r}
#Question 22

Event_Region %>% 
  select(name, sex, age, team, games, 
         city, sport, event, medal, region) %>% 
  arrange(desc(age)) %>% 
  filter(medal == "Gold", age == 64)
```

**Question 23:**
Who are the top 10 athletes who have won the most gold medals?
```{r}
#Question 23

Event_Region %>% 
  select(name, total_gold_medals = medal, team) %>% 
  filter(total_gold_medals == "Gold") %>% 
  aggregate(total_gold_medals ~ name + team, FUN = length) %>% 
  arrange(desc(total_gold_medals)) %>% 
  slice(1:10)
```

**Question 24:**
Get the top 10 athletes by number of medals earned (gold, silver, or bronze).
```{r}
#Question 24

Event_Region %>% 
  select(name, total_medals = medal, team) %>% 
  aggregate(total_medals ~ name + team, FUN = length) %>% 
  arrange(desc(total_medals)) %>% 
  slice(1:10)
```

**Question 25:**
What age group participated in the game the most?
```{r}
#Question 25

Event_Region %>% 
  group_by(age_group = age) %>% 
  summarise(count=n()) %>%
  na.omit() %>%
  arrange(desc(count))
```

The most age group is 23 years of age with 21,848 participant
